---
title: "Lab #5 MA 350-01, Spring 2020"
author: "Carlin, Jonathan"
output: html_notebook
---

# Git and GitHub

## Problem 0:

### URL: https://github.com/JonC0095/MA350-Lab5


# Finding Outlier Using Tukey's Rule-of-Thumb

## Problem 1:

```{r}
x <-c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
sort.x <- sort(x)

q1.x <- quantile(x,0.25)[[1]]
q3.x <- quantile(x,0.75)[[1]]
IQR.x <- (q3.x - q1.x)

x
sort.x
length(x)
q1.x
q3.x
IQR.x
```

I have sorted x and found the length of x to be 12. This indicates to me that the first quartile should be the 3rd term in and that third quartile should be the 9th term. We should also note that 28.7 is a possible outlier and might skew this. Thus, q1.x produced -4.525 which seems reasonably close to the assumption, and q3.x produced 4.9, which is the correct value. 4.9-(-4.525)= 9.425 and so the value produced by IQR.x is correct.


## Problem 2:
```{r}

quartiles <- function(x){
  
  length(x)
  
  quantiles.x <- quantile(x, probs=c(0.25, 0.75))
  
  quant.v <- c(quantiles.x[[1]],quantiles.x[[2]], quantiles.x[[2]]-quantiles.x[[1]])
  
  return(quant.v)
  
}

quartiles(x)

```

The results, (-4.525), 4.9, & 9.452, are the same to those found in problem 1 for the first quantile, third quantile, & IQR, respectively.

## Problem 3:
```{r}

#low.outlier.max <- q1.x - (1.5 * IQR.x)
#high.outlier.min <- q3.x + (1.5 * IQR.x)
#needed to come back and fix this because it is hard coded in and not based off x

low.outlier.max <- quartiles(x)[1] - (1.5 * quartiles(x)[3])
high.outlier.min <- quartiles(x)[2] + (1.5 * quartiles(x)[3])

i <- 1

for(i in 1:(length(x)-1)){
  
  if(x[i] < low.outlier.max){
    low.outlier.x <- x[i]
    i <- i + 1
  }
  else if(x[i] > high.outlier.min){
    high.outlier.x <- x[i]
    i <- i + 1
  }
  else{
    i <- i + 1
  }

}


if(exists("low.outlier.x")== TRUE){ print("Lower outlier(s):")
  show(low.outlier.x) }
if(exists("high.outlier.x")== TRUE){ print("Upper outlier(s):")
  show(high.outlier.x)  }


```

We can see that 28.7 is an upper outlier, this agrees with problem 1 answer.

## Problem 4:

I see I misunderstood problem 3; I will adapt to fit problem 4.

```{r}

tukey.outlier <- function(x){
  
bool.outlier <- logical(length(x))


i <- 1

for(i in 1:(length(x)-1)){
  
  if(x[i] < low.outlier.max){ bool.outlier[i]=TRUE
    i <- i + 1 }
  
  else if(x[i] > high.outlier.min){ bool.outlier[i]=TRUE
    i <- i + 1  }
  
  else{ bool.outlier[i]=FALSE
    i <- i + 1  }
}


return(bool.outlier)
  
}

x
tukey.outlier(x)
```

we can see, from problem 3 and 4 that 28.7 is the only outlier. 
 
## Problem 5:
```{r}

test.tukey.outlier <- function( FUN = tukey.outlier){
  
  x <-c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
  bool.x <-c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE)
  scalar.x <- ((0.1)*x)
  
  
  i <- 1
  
  for(i in 1:(length(x)-1)){
    
  if( (tukey.outlier(x)[i] == bool.x[i]) & (tukey.outlier(-x)[i] == bool.x[i]) ){
    
    outcome.x <- TRUE
    outcome_x <- TRUE
    
    if( (tukey.outlier(scalar.x)[i] == bool.x[i]) ){
      outcome.scalar <- TRUE
    }else{ 
      return(FALSE)  
    }
    
    i <- i + 1 
  } 
  
    else{ 
      return(FALSE)  
    }
    
  }
  
  outcome <- c(outcome.x, outcome_x,outcome.scalar)
  
  
  return(outcome)
  
}

test.tukey.outlier(x)

```

## Problem 6: 

An outlier in the data values is -28.7; changing the sign does not change their relative distance because we are affecting the whole set.


## Problem 7 & 8:

```{r}

tukey.outlier(x)
test.tukey.outlier(x)

```

## Problem 9:

An outlier in the data values is (0.1)(28.7)=2.87; multiplying by a scalar does not change their relative distance because we are affecting the whole set.

## Problem 10 & 11:

I began to modify test.tukey.outlier to include 0.1*x as an additiional test case; however, I ran into issues.
To start, I realized I had code that would not be altered by the change in x but rather were hard coded in.
Once fixed, I realized That I had to be more careful with my usage of & vs &&; however I am stuck now..
When multiplying by a scalar I have an issue.







