---
title: "Lab #5 MA 350-01, Spring 2020"
author: "Carlin, Jonathan"
output: html_notebook
---

# Git and GitHub

## Problem 0:

### URL: https://github.com/JonC0095/MA350-Lab5


# Finding Outlier Using Tukey's Rule-of-Thumb

## Problem 1:

```{r}
x <-c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
sort.x <- sort(x)

q1.x <- quantile(x,0.25)
q3.x <- quantile(x,0.75)
IQR.x <- q3.x - q1.x

sort.x
length(x)
q1.x[[1]]
q3.x[[1]]
IQR.x[[1]]
```

I have sorted x and found the length of x to be 12. This indicates to me that the first quartile should be the 3rd term in and that third quartile should be the 9th term. We should also note that 28.7 is a possible outlier and might skew this. Thus, q1.x produced -4.525 which seems reasonably close to the assumption, and q3.x produced 4.9, which is the correct value. 4.9-(-4.525)= 9.425 and so the value produced by IQR.x is correct.


## Problem 2:
```{r}

quartiles <- function(x){
  
  quantiles.x <- quantile(x, probs=c(0.25, 0.75))
  
  quant.v <- c(quantiles.x[[1]],quantiles.x[[2]], quantiles.x[[2]]-quantiles.x[[1]])
  
  return(quant.v)
  
}

quartiles(x)
```

The results, (-4.525), 4.9, & 9.452, are the same to those found in problem 1 for the first quantile, third quantile, & IQR, respectively.

## Problem 3:
```{r}

Outlier.min

Outlier.max

quartiles(x)

```



